---
title: "Fig_3_border_AOR_version"
author: "Claire Morton"
date: "2026-06-12"
output: html_document
---

We want to get the before and after arrest rate gap for border and non-border AORs.
Border AORs: San Diego, Phoenix, El Paso, San Antonio, Harlingen
Non-border AORs: All others
This analysis will use the AOR arrest category, which doesn't always match with state in a way that makes sense.

## TODO: Number of cases where state doesn't match city AOR

```{r setup}
library(ggplot2)
library(dplyr)
library(fixest)
library(forcats)
```

## Functions
```{r}
aggregate_data <- function(data, 
                           cols_to_keep, 
                           pop_col,
                           agg_noneuca = TRUE){
  
  
  pop_cols <- c("pop_ACS", "pop_MPI")
  
  data_pop <- data %>% 
    select(-setdiff(pop_cols, pop_col)) %>%
    rename("pop" = !!sym(pop_col))
  
  raw_pop <- data_pop %>%
    group_by(aor, region) %>%
    summarize(pop = first(pop), .groups = "drop")
  
  # aggregate regions if reuqested
  if(agg_noneuca){
    data_pop <- data_pop %>% 
      mutate(region = ifelse(region == "EUCA", "EUCA", "nonEUCA"))
    
    raw_pop <- raw_pop %>% 
      mutate(region = ifelse(region == "EUCA", "EUCA", "nonEUCA")) %>%
      group_by(aor, region) %>%
      summarize(pop = sum(pop), .groups = "drop")
  }
  
  # checksum value
  checksum <- data_pop %>%
    group_by(aor, region, pop) %>%
    slice_head(n=1) %>%
    ungroup() %>%
    group_by(region) %>%
    summarize(pop_checksum = sum(pop), .groups = "drop")
  
  # find geographic keys
  geo_keys <- intersect(cols_to_keep, c("aor", "region"))
  
  # population map
  pop_map <- data_pop %>%
    group_by(aor, region, pop) %>%
    slice_head(n = 1) %>%
    ungroup() %>%
    group_by_at(geo_keys) %>%
    summarize(pop = sum(pop), .groups = "drop")
  
  # aggregate arrests, join population back in
  data_pop <- data_pop %>%
    group_by_at(cols_to_keep) %>%
    summarize(n_arrests = sum(n_arrests, na.rm = TRUE), .groups = "drop") %>%
    left_join(pop_map, by = geo_keys) %>%
    mutate(ratio = n_arrests/pop*100000,
           log_pop = log(pop))
  
  # checksum
  total_aggregated <- data_pop %>%
    group_by_at(geo_keys) %>%
    summarize(pop = first(pop), .groups = "drop") %>%
    pull(pop) %>%
    sum(na.rm = TRUE)
  total_checksum <- sum(checksum$pop_checksum, na.rm = TRUE)
  if (!(dplyr::near(total_checksum, total_aggregated))) {
    warning("Checksum Failed!")
  }
  
  return(data_pop)
}

jackknife_supt <- function(fe_fit, model_fml, data, model_type = c("linear", "poisson"), extra_fml = NULL, do_se = TRUE) {
  model_type <- match.arg(model_type)
  coef0      <- coef(fe_fit)
  aors     <- unique(data$aor)
  n_aors   <- length(aors)

  # Event-time labels (excludes reference period -1)
  windows <- c(1)

  if (do_se) {
    # G x K matrix of leave-one-aor-out coefficients
    jackknife_est_matrix <- t(sapply(aors, function(dropped_aor) {
      sub_data <- subset(data, aor != dropped_aor)

      if (model_type == "poisson") {
        coef(fepois(model_fml, data = sub_data, offset = extra_fml, cluster = ~aor))
      } else {
        coef(feols(model_fml, data = sub_data, weights = extra_fml, cluster = ~aor))
      }
    }))

    # TODO: Is this correct?
    jackknife_mean <- mean(jackknife_est_matrix)
    jackknife_var <- (n_aors -1 )/(n_aors) * sum( (jackknife_est_matrix - jackknife_mean)^2)
    jackknife_se <- sqrt(jackknife_var)
    
    # TODO: critical value?
    critical_value = 1.96
    
    # Apply exponential transformation if Poisson
    if (model_type == "poisson") {
      table_out <- tibble(window = windows,
                     est    = exp(coef0), se = jackknife_se,
                     lower  = exp(coef0 - critical_value * se),
                     upper  = exp(coef0 + critical_value * se))
    } else {
      table_out <- tibble(window = windows,
                     est    = coef0, 
                     se = jackknife_se,
                     lower  = coef0 - critical_value * se,
                     upper  = coef0 + critical_value * se)
    }
  } else {
    est_val <- if (model_type == "poisson") exp(coef0) else coef0
    table_out    <- tibble(window = windows, est = est_val, se = 0, lower = est_val, upper = est_val)
  }

  # Add the reference period (-1 point) dynamically
  table_out
}

fit_diffindiff <- function(data, model_fml, extra_fml, model_type = c("linear", "poisson"), do_se = TRUE) {
  model_type  <- match.arg(model_type)
  region_vals <- setdiff(unique(data$region), "EUCA")
  region_coefs <- list()
  
  for (cur_region in region_vals) {
    data_filtered <- data %>% filter(region %in% c("EUCA", cur_region))
    
    if (model_type == "poisson") {
      mod <- fepois(model_fml, data_filtered, offset = extra_fml)
    } else {
      mod <- feols(model_fml, data_filtered, weights = extra_fml, cluster = ~aor)
    }
    
    region_coefs[[cur_region]] <- jackknife_supt(
      fe_fit    = mod, 
      model_fml = model_fml, 
      data      = data_filtered, 
      model_type = model_type, 
      extra_fml = extra_fml, 
      do_se     = do_se
    ) %>%
      mutate(region = cur_region)
  }
  
  return(bind_rows(region_coefs))
}

event_study_plot <- function(estimates,
                             pal,
                             baseline = 0) {
  estimates |>
    ggplot(aes(x = est, y = group, color = aor_group, group = aor_group)) +
    geom_vline(xintercept = baseline, linetype = "dashed", color = "grey50") +
    geom_point(size = 2.5) +
    geom_errorbar(aes(xmin = lower, xmax = upper)) +
    geom_line(linewidth = 0.8) +
    labs(x = "Difference in Arrests per 100k\nrelative to EUCA and pre-Trump period",
         y = "Group") +
    theme_minimal()+
    scale_color_manual(values = pal)
}

arrange_figure_3 <- function(trajectory, model_fml, wt_fml, model_type, do_se = T){
  model_list = list()
  for (cur_group in unique(trajectory$group)){
    trajectory_filtered <- trajectory %>% filter(group == cur_group)
    # once we have just that group, run the usual diff in diff with the SEs
    res <- fit_diffindiff(
        trajectory_filtered,
        model_fml,
        wt_fml,
        model_type = model_type,
        do_se = do_se
      ) %>%
        mutate(aor_group = cur_group)
      
    model_list[[cur_group]] = res
  }
  
  model_list = bind_rows(model_list) %>%
    mutate(group = aor_group)
  event_study_plot(model_list, pal = pal_group, baseline = 1) +
      labs(color = "AOR Group")
}
```

## Figures

Want to plot the before vs after EUCA vs non-EUCA difference, use AOR as the standard error jackknife instead of state.
So just two bars, and 0.

```{r setup_data, include = F}
#### CONSTANTS ####
model_fml <- as.formula(
  paste0("ratio ~ i(window, region, ref = -1, ref2 = 'EUCA') | region + window")
)
wt_fml <- as.formula("~ pop")
model_type = "linear"

pois_model_fml <- as.formula(
  paste0("n_arrests ~ i(window, region, ref = -1, ref2 = 'EUCA') | region + window")
)
offset_fml <- as.formula("~log_pop")

groups = c("Border", "Non-border")
pal_group <- c(setNames(scales::brewer_pal(palette = "Dark2")(4),groups))
```


```{r, main_figs, echo = F}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv") 

trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_MPI",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000,
         log_pop = log(pop))

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
  
model_type = "poisson"
mult_denom <- arrange_figure_3(trajectory, pois_model_fml, offset_fml, model_type, do_se = T)+
  ggtitle("Multiplicative", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio/pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect/euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

mult_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity") +
  geom_vline(xintercept = 1)+
  ggtitle("Multiplicative")
```

```{r, main_figs, echo = F}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv") 

trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_ACS",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000)

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
model_type = "linear"
acs_denom <- arrange_figure_3(trajectory, model_fml, wt_fml, model_type, do_se = T)+
  ggtitle("ACS Denom", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio - pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect-euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

acs_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity") +
  ggtitle("ACS")
```

```{r}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv") 

trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_MPI",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000)

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
  
mpi_denom <- arrange_figure_3(trajectory, model_fml, wt_fml, model_type, do_se = T)+
  ggtitle("MPI Denom", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio - pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect-euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

mpi_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity")+
  ggtitle("MPI")
```

```{r}
ggarrange(mpi_denom, acs_denom)
ggarrange(mpi_effects, acs_effects)
ggarrange(mult_denom, mult_effects)
```


```{r, mult_figs_ca, echo = F}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv") %>%
  filter(method == "CA")

model_type = "poisson"
 
trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_MPI",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000,
         log_pop = log(pop))

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
  
model_type = "poisson"
mult_denom <- arrange_figure_3(trajectory, pois_model_fml, offset_fml, model_type, do_se = T)+
  ggtitle("Multiplicative Community Arrests", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio/pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect/euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

mult_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity") +
  geom_vline(xintercept = 1)+
  ggtitle("Multiplicative Community Arrests")
```

```{r, main_figs_ca, echo = F}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv")  %>%
  filter(method == "CA")

model_type = "linear"

trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_ACS",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000)

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
  
acs_denom <- arrange_figure_3(trajectory, model_fml, wt_fml, model_type, do_se = T)+
  ggtitle("ACS Denom Community Arrests", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio - pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect-euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

acs_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity") +
  ggtitle("ACS Community Arrests")
```

```{r, figs_ca, echo = F}
trajectory <- read_csv("data/glm_trajectory_type_threatlevel_convicted_ACS_MPI_AOR.csv")  %>%
  filter(method == "CA")

trajectory <- aggregate_data(trajectory, 
                             c("region", "aor", "window"), 
                             pop_col = "pop_MPI",
                             agg_noneuca = T) %>%
  mutate(window = ifelse(window >= 0, 1, -1)) %>%
  select(-c(ratio, log_pop)) %>%
  group_by(region, aor, window, pop) %>%
  summarize(n_arrests = sum(n_arrests)) %>%
  mutate(ratio = n_arrests/pop*100000)

trajectory <- trajectory %>%
  mutate(group = ifelse(aor %in% c("San Diego", "Phoenix", "El Paso", "San Antonio", "Harlingen"), "Border", "Non-border"))
  
mpi_denom <- arrange_figure_3(trajectory, model_fml, wt_fml, model_type, do_se = T)+
  ggtitle("MPI Denom Community Arrests", subtitle = "Border = San Diego, Phoenix, El Paso, San Antonio, Harlingen")

effects <- trajectory %>%
  ungroup() %>%
  filter(window == -1) %>%
  select(region, aor, ratio) %>%
  rename("pre_ratio" = ratio) %>%
  merge(trajectory) %>%
  filter(window == 1) %>%
  mutate(effect = ratio - pre_ratio) %>%
  arrange(effect)

diffindiff_effects <- effects %>%
  filter(region == "EUCA") %>%
  select(aor, effect) %>%
  rename("euca_effect" = effect) %>%
  merge(effects) %>%
  filter(region == "nonEUCA") %>%
  mutate(effect = effect-euca_effect) %>%
  mutate(aor = fct_reorder(aor, desc(effect)))

mpi_effects <- ggplot(diffindiff_effects, aes(y = aor, x = effect, fill = group)) +
  geom_bar(stat = "identity")+
  ggtitle("MPI Community Arrests")
```

```{r}
ggarrange(mpi_denom, acs_denom)
ggarrange(mpi_effects, acs_effects)
ggarrange(mult_denom, mult_effects)
```