---
title: "Paper_Numbers"
author: "Claire Morton"
date: "2026-06-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(readr)
library(haven)
library(lubridate)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(arrow)
library(scales)
```

Our primary data source is administrative from the Deportation Data Project (DDP) obtained through FOIA litigation, on the universe of ICE Enforcement and Removal Operations from [DATE to DATE] (n=XXX). 
```{r }
raw_data <- read_parquet("../data/arrests-latest.parquet")

# plot hist of apprehension dates
hist(as.Date(raw_data$apprehension_date), breaks = 20)

# sum restricted number of total arrests 
raw_data <- raw_data %>%
  filter((as.Date(apprehension_date) >= as.Date("2022-10-01")) &
          (as.Date(apprehension_date) <= as.Date("2026-03-10")))

nrow(raw_data)
  
```
Dates run from 10-01-2022 to 03-10-2026
We have a total of 713464 arrests that occur in these dates (2 were outside of them in the raw data)

Restrictions of the data
We restrict the sample to exclude minors (n = ), arrestees with unknown state of arrest (n = ); see SI.
```{r }
MCA = c('MEXICO','GUATEMALA','EL SALVADOR','HONDURAS','NICARAGUA','COSTA RICA','PANAMA','BELIZE')

CARIBBEAN = c('CUBA','DOMINICAN REPUBLIC','HAITI','JAMAICA','TRINIDAD AND TOBAGO','BAHAMAS','BARBADOS',
  'ANTIGUA AND BARBUDA','SAINT LUCIA','GRENADA','DOMINICA','SAINT VINCENT AND THE GRENADINES',
  'SAINT KITTS AND NEVIS','ST. LUCIA','ST. KITTS-NEVIS','ST. VINCENT-GRENADINES','ANTIGUA-BARBUDA',
  'BRITISH VIRGIN ISLANDS','TURKS AND CAICOS ISLANDS','BERMUDA','NETHERLANDS ANTILLES','GUADELOUPE',
  'CURACAO','ARUBA','ANGUILLA','MONTSERRAT','CAYMAN ISLANDS','SINT MAARTEN(DUTCH)','SINT EUSTATIUS'
)
SA = c('BRAZIL','COLOMBIA','VENEZUELA','ECUADOR','PERU','ARGENTINA','BOLIVIA','CHILE','PARAGUAY',
  'URUGUAY','GUYANA','SURINAME','FRENCH GUIANA')
ASIA = c(
  'CHINA','INDIA','PHILIPPINES','VIETNAM','SOUTH KOREA','JAPAN','BANGLADESH','PAKISTAN',
  'INDONESIA','MALAYSIA','THAILAND','SRI LANKA','NEPAL','BURMA','MYANMAR','CAMBODIA','LAOS',
  'SINGAPORE','MONGOLIA','TAIWAN','AFGHANISTAN','UZBEKISTAN','KAZAKHSTAN','KYRGYZSTAN','TAJIKISTAN',
  'TURKMENISTAN','NORTH KOREA','BHUTAN','MALDIVES','BRUNEI','IRAN','IRAQ','SYRIA','LEBANON','JORDAN',
  'ISRAEL','YEMEN','SAUDI ARABIA','TURKEY','ARMENIA','AZERBAIJAN','GEORGIA','KUWAIT','QATAR',
  'BAHRAIN','UNITED ARAB EMIRATES','OMAN','PALESTINE',
  # DDP-specific spellings:
  'CHINA, PEOPLES REPUBLIC OF','TURKIYE','KOREA','HONG KONG','MACAU','EAST TIMOR',
  'PALESTINE BORN BEFORE 1948'
)
AFRICA = c(
  'NIGERIA','ETHIOPIA','EGYPT','KENYA','GHANA','SOMALIA','SUDAN','SOUTH SUDAN','SOUTH AFRICA',
  'SENEGAL','CAMEROON','LIBERIA','SIERRA LEONE','ERITREA','IVORY COAST',"COTE D'IVOIRE",'MALI',
  'MOROCCO','TUNISIA','ALGERIA','UGANDA','TANZANIA','ZIMBABWE','RWANDA','BURKINA FASO','CHAD',
  'NIGER','GAMBIA','GUINEA','TOGO','BENIN','MAURITANIA','ANGOLA','MOZAMBIQUE','ZAMBIA','MALAWI',
  'MADAGASCAR','BOTSWANA','NAMIBIA','LESOTHO','CONGO','DEMOCRATIC REPUBLIC OF THE CONGO',
  'CENTRAL AFRICAN REPUBLIC','EQUATORIAL GUINEA','GABON','COMOROS','CAPE VERDE','DJIBOUTI',
  'BURUNDI','LIBYA',
  # DDP-specific spellings:
  'DEM REP OF THE CONGO','GUINEA-BISSAU','MAURITIUS','SAO TOME AND PRINCIPE','ESWATINI'
)
EUCA = c(
  'CANADA','UNITED KINGDOM','IRELAND','GERMANY','FRANCE','ITALY','SPAIN','PORTUGAL','GREECE',
  'RUSSIA','UKRAINE','POLAND','ROMANIA','HUNGARY','CZECH REPUBLIC','SLOVAKIA','BULGARIA','SERBIA',
  'CROATIA','SLOVENIA','MACEDONIA','NORTH MACEDONIA','ALBANIA','BOSNIA AND HERZEGOVINA','BELGIUM',
  'NETHERLANDS','SWEDEN','NORWAY','DENMARK','FINLAND','ICELAND','SWITZERLAND','AUSTRIA',
  'LUXEMBOURG','CYPRUS','MALTA','ESTONIA','LATVIA','LITHUANIA','BELARUS','MOLDOVA','MONTENEGRO',
  'KOSOVO','AUSTRALIA','NEW ZEALAND','FIJI','SAMOA','TONGA','PAPUA NEW GUINEA',
  # DDP-specific spellings:
  'BOSNIA-HERZEGOVINA','USSR','YUGOSLAVIA','SERBIA AND MONTENEGRO','CZECHOSLOVAKIA','ANDORRA',
  'MONACO','MICRONESIA, FEDERATED STATES OF','MARSHALL ISLANDS','PALAU','FRENCH POLYNESIA'
)


get_region <- function(country){
  if (country %in% MCA){
    return("MCA")
  } else if (country %in% CARIBBEAN){
    return("CARIBBEAN")
  } else if (country %in% SA){
    return("SA")
  } else if (country %in% ASIA){
    return("ASIA")
  } else if (country %in% AFRICA){
    return("AFRICA")
  } else if (country %in% EUCA){
    return("EUCA")
  } else{
    return(NA)
  }
}

get_method <- function(method){
  if (method %in% LEA_METHODS){
    return("LEA")
  } else if (method %in% CA_METHODS){
    return("CA")
  } else{
    return("OTHER")
  }
}

LEA_METHODS = c('287(g) Program','Anti-Smuggling','CAP Federal Incarceration','CAP Local Incarceration',
  'CAP State Incarceration','Criminal Alien Program','ERO Reprocessed Arrest',
  'Law Enforcement Agency Response Unit','Organized Crime Drug Enforcement Task Force',
  'Other Agency (turned over to INS)','Other Task Force','Probation and Parole',
  'Custodial Arrest','Patrol Border','Patrol Interior')

CA_METHODS = c('Located','Non-Custodial Arrest','Worksite Enforcement')

STATES = c(toupper(state.name), "DISTRICT OF COLUMBIA")
```

```{r}
raw_arrests <- read_parquet("../data/arrests-latest.parquet") %>%
  rename("ApprehensionDate" = "apprehension_date", 
         "ApprehensionAOR" = "apprehension_aor",
         "ApprehensionMethod" = "apprehension_method",
         "ApprehensionCriminality" = "apprehension_criminality"
         ) %>%
  select("ApprehensionDate","ApprehensionAOR",
         "ApprehensionMethod","ApprehensionCriminality","apprehension_state",
         "citizenship_country", "gender", "birth_year", "case_threat_level")

raw_arrests <- raw_arrests %>%
  mutate(aor_short = sub(" Area of Responsibility", "", ApprehensionAOR))

raw_arrests[is.na(raw_arrests$apprehension_state),]$apprehension_state = ""

raw_arrests$state_clean = raw_arrests$apprehension_state


STATES = c(toupper(state.name), "DISTRICT OF COLUMBIA")
raw_arrests[!(raw_arrests$state_clean %in% STATES),]$state_clean = ""

raw_arrests <- raw_arrests |> 
  mutate(state_clean = case_when(
    aor_short == "Phoenix" ~ "ARIZONA",
    aor_short %in% c("Buffalo", "New York City") ~ "NEW YORK",
    aor_short %in% c("San Diego", "Los Angeles") ~ "CALIFORNIA",
    aor_short %in% c("Houston", "Harlingen", "San Antonio", "Dallas") ~ "TEXAS",
    aor_short == "Miami" ~ "FLORIDA",
    TRUE ~ state_clean 
  ))

raw_arrests[is.na(raw_arrests$aor_short),]$aor_short = ""
raw_arrests[raw_arrests$aor_short == "HQ",]$aor_short = ""

raw_arrests <- raw_arrests |>
  mutate(aor_short = case_when(
    aor_short == "" & state_clean %in% c("ALASKA", "WASHINGTON", "OREGON") ~ "Seattle",
    aor_short == "" & state_clean %in% c("HAWAII") ~ "San Francisco",
    aor_short == "" & state_clean %in% c("UTAH", "NEVADA", "IDAHO", "MONTANA") ~ "Salt Lake City",
    aor_short == "" & state_clean %in% c("ARIZONA") ~ "Phoenix",
    aor_short == "" & state_clean %in% c("NEW MEXICO") ~ "El Paso",
    aor_short == "" & state_clean %in% c("OKLAHOMA") ~ "Dallas",
    aor_short == "" & state_clean %in% c("WYOMING", "COLORADO") ~ "Denver",
    aor_short == "" & state_clean %in% c("UTAH", "NEVADA", "IDAHO", "MONTANA") ~ "Salt Lake City",
    aor_short == "" & state_clean %in% c("MINNESOTA", "IOWA", "NEBRASKA", "SOUTH DAKOTA", "NORTH DAKOTA") ~ "St. Paul",
    aor_short == "" & state_clean %in% c("WISCONSIN", "ILLINOIS", "INDIANA", "KENTUCKY", "MISSOURI", "KANSAS") ~ "Chicago",
    aor_short == "" & state_clean %in% c("ARKANSAS", "LOUISIANA", "TENNESSEE", "MISSISSIPPI", "ALABAMA") ~ "New Orleans",
    aor_short == "" & state_clean %in% c("FLORIDA") ~ "Miami",
    aor_short == "" & state_clean %in% c("GEORGIA", "SOUTH CAROLINA", "NORTH CAROLINA") ~ "Atlanta",
    aor_short == "" & state_clean %in% c("VIRGINIA", "DISTRICT OF COLUMBIA") ~ "Washington",
    aor_short == "" & state_clean %in% c("MICHIGAN", "OHIO") ~ "Detroit",
    aor_short == "" & state_clean %in% c("MARYLAND") ~ "Baltimore",
    aor_short == "" & state_clean %in% c("NEW JERSEY") ~ "Newark",
    aor_short == "" & state_clean %in% c("DELAWARE", "PENNSYLVANIA", "WEST VIRGINIA") ~ "Philadelphia",
    aor_short == "" & state_clean %in% c("VERMONT", "NEW HAMPSHIRE", "MAINE", "MASSACHUSETTS", "CONNECTICUT", "RHODE ISLAND") ~ "Boston",
    TRUE ~ aor_short
  ))
```

```{r}
# obtain count of minors:
print("Minors:")
sum(((year(ymd(raw_arrests$ApprehensionDate)) - raw_arrests$birth_year) < 18), na.rm = T)/nrow(raw_arrests) *100

arrests <- raw_arrests %>%
  filter((as.Date(ApprehensionDate) >= as.Date("2023-11-27")) &
        (as.Date(ApprehensionDate) <= as.Date("2026-03-10"))) %>%
  filter((year(ymd(ApprehensionDate)) - birth_year) >= 18) %>% # filter out children
  select(c(ApprehensionDate, ApprehensionMethod, 
           ApprehensionCriminality, case_threat_level,
           state_clean, aor_short, citizenship_country)) %>%
   mutate(citizenship_region = unlist(lapply(citizenship_country, get_region)))

# missing location information
print("Missing location information")
sum((raw_arrests$state_clean == "") | (raw_arrests$aor_short == ""))/nrow(raw_arrests) *100

arrests <- arrests %>%
  filter(state_clean != "" & aor_short != "")

print("Region of Origin Unknown:")
sum(is.na(arrests$citizenship_region))/nrow(raw_arrests) *100

arrests <- arrests %>%
  filter(!is.na(citizenship_region))
```

We test each of these claims using data on more than half a million administrative ICE arrests from [DATE] to March 10, 2026 (the most recent date currently available) from the Deportation Data Project (CITE). 
To do so, we first temporally aggregate arrests into 14 60-day periods, anchored at January 20, 2025. Our study period thus ranges from TKTK to TKTK pre-treatment, and from TKTK to TKTK post-treatment (n = TKTK). 
```{r}
INAUG_DATE = "2025-01-20"
WINDOW_DAYS = 60
TOTAL_DAYS_EACH_SIDE = 7*WINDOW_DAYS

assign_window <- function(dates){
  days_from_inaug = as.numeric(as.Date(dates) - as.Date(INAUG_DATE))
  window_assignment = ifelse(days_from_inaug < 0, -((-days_from_inaug - 1) %/% WINDOW_DAYS + 1),
                             days_from_inaug %/% WINDOW_DAYS)
  window_assignment = ifelse(abs(days_from_inaug) > TOTAL_DAYS_EACH_SIDE, NA, window_assignment)
  return(window_assignment)
}

arrests$window <- assign_window(arrests$ApprehensionDate)

print("Start of first window:")
as.Date("2025-01-20")-60*7

print("End of final window:")
as.Date("2025-01-20")+60*7-1

print("Max date available in data:")
max((arrests %>% filter(window == max(arrests$window, na.rm = T)) )$ApprehensionDate)

print("Outside of window:")
nrow(arrests %>% filter(is.na(window)))
nas <- arrests %>%
  filter(is.na(window))

print("Inside of window:")
arrests <- arrests %>%
  filter(!is.na(window))
nrow(arrests)

summary(as.factor(do.call(rbind, lapply(arrests$ApprehensionMethod, get_method))))
```


```{r}
# load raw arrests data to get the other variables
print(paste0("Number of arrests: ", nrow(raw_arrests)))

raw_arrests <- raw_arrests %>%
  filter((as.Date(ApprehensionDate) >= as.Date("2023-11-27")) &
        (as.Date(ApprehensionDate) <= as.Date("2026-03-10")))
print(paste0("Filtering to dates of interest: ", nrow(raw_arrests)))

raw_arrests <- raw_arrests %>%
  filter(((year(ymd(ApprehensionDate)) - birth_year) >= 18))
print(paste0("Remove minors: ", nrow(raw_arrests)))

raw_arrests <- raw_arrests %>%
  filter(state_clean != "")
print(paste0("Remove no state: ", nrow(raw_arrests)))

raw_arrests <- raw_arrests %>%
  filter(aor_short != "")
print(paste0("Remove no aor: ", nrow(raw_arrests)))

raw_arrests_no_region <- raw_arrests %>%
   mutate(citizenship_region = unlist(lapply(citizenship_country, get_region))) %>%
  filter(is.na(citizenship_region))

raw_arrests <- raw_arrests %>%
   mutate(citizenship_region = unlist(lapply(citizenship_country, get_region))) %>%
  filter(!is.na(citizenship_region))

print(paste0("Require valid region: ", nrow(raw_arrests)))
```

Figure out how many AORs mismatch their states
```{r}
valid_state_aor <- raw_arrests %>%
  rename("aor" = "aor_short",
         "state" = "state_clean") %>%
  select(aor, state)

valid_state_aor <- valid_state_aor %>%
  filter((aor == "Seattle" & state %in% c("ALASKA", "WASHINGTON", "OREGON")) |
           (aor == "San Francisco" & state %in% c("CALIFORNIA", "HAWAII")) |
           (aor == "San Diego" & state == "CALIFORNIA") |
           (aor == "Los Angeles" & state == "CALIFORNIA") |
           (aor == "Salt Lake City" & state %in% c("UTAH", "NEVADA", "IDAHO", "MONTANA")) |
           (aor == "Phoenix" & state == "ARIZONA") |
           (aor == "El Paso" & state %in% c("TEXAS", "NEW MEXICO")) |
           (aor == "San Antonio" & state == "TEXAS") |
           (aor == "Harlingen" & state == "TEXAS") |
           (aor == "Houston" & state == "TEXAS") |
           (aor == "Dallas" & state %in% c("TEXAS", "OKLAHOMA")) |
           (aor == "Denver" & state %in% c("WYOMING", "COLORADO")) |
           (aor == "St. Paul" & state %in% c("MINNESOTA", "IOWA", "NEBRASKA", "SOUTH DAKOTA", "NORTH DAKOTA")) |
           (aor == "Chicago" & state %in% c("WISCONSIN", "ILLINOIS", "INDIANA", "KENTUCKY", "MISSOURI", "KANSAS"))|
           (aor == "New Orleans" & state %in% c("ARKANSAS", "LOUISIANA", "TENNESSEE", "MISSISSIPPI", "ALABAMA")) |
           (aor == "Miami" & state == "FLORIDA") |
           (aor == "Atlanta" & state %in% c("GEORGIA", "SOUTH CAROLINA", "NORTH CAROLINA")) |
           (aor == "Washington" & state %in% c("VIRGINIA", "DISTRICT OF COLUMBIA")) |
           (aor == "Detroit" & state %in% c("MICHIGAN", "OHIO")) |
           (aor == "Baltimore" & state == "MARYLAND") |
           (aor == "Newark" & state == "NEW JERSEY") |
           (aor == "Philadelphia" & state %in% c("DELAWARE", "PENNSYLVANIA", "WEST VIRGINIA")) |
           (aor == "New York City" & state == "NEW YORK") |
           (aor == "Buffalo" & state == "NEW YORK") |
           (aor == "Boston" & state %in% c("VERMONT", "NEW HAMPSHIRE", "MAINE", "MASSACHUSETTS", "CONNECTICUT", "RHODE ISLAND")))


invalid_state_aor <- raw_arrests %>%
  rename("aor" = "aor_short",
         "state" = "state_clean")

invalid_state_aor <- invalid_state_aor %>%
  filter(!((aor == "Seattle" & state %in% c("ALASKA", "WASHINGTON", "OREGON")) |
           (aor == "San Francisco" & state %in% c("CALIFORNIA", "HAWAII")) |
           (aor == "San Diego" & state == "CALIFORNIA") |
           (aor == "Los Angeles" & state == "CALIFORNIA") |
           (aor == "Salt Lake City" & state %in% c("UTAH", "NEVADA", "IDAHO", "MONTANA")) |
           (aor == "Phoenix" & state == "ARIZONA") |
           (aor == "El Paso" & state %in% c("TEXAS", "NEW MEXICO")) |
           (aor == "San Antonio" & state == "TEXAS") |
           (aor == "Harlingen" & state == "TEXAS") |
           (aor == "Houston" & state == "TEXAS") |
           (aor == "Dallas" & state %in% c("TEXAS", "OKLAHOMA")) |
           (aor == "Denver" & state %in% c("WYOMING", "COLORADO")) |
           (aor == "St. Paul" & state %in% c("MINNESOTA", "IOWA", "NEBRASKA", "SOUTH DAKOTA", "NORTH DAKOTA")) |
           (aor == "Chicago" & state %in% c("WISCONSIN", "ILLINOIS", "INDIANA", "KENTUCKY", "MISSOURI", "KANSAS"))|
           (aor == "New Orleans" & state %in% c("ARKANSAS", "LOUISIANA", "TENNESSEE", "MISSISSIPPI", "ALABAMA")) |
           (aor == "Miami" & state == "FLORIDA") |
           (aor == "Atlanta" & state %in% c("GEORGIA", "SOUTH CAROLINA", "NORTH CAROLINA")) |
           (aor == "Washington" & state %in% c("VIRGINIA", "DISTRICT OF COLUMBIA")) |
           (aor == "Detroit" & state %in% c("MICHIGAN", "OHIO")) |
           (aor == "Baltimore" & state == "MARYLAND") |
           (aor == "Newark" & state == "NEW JERSEY") |
           (aor == "Philadelphia" & state %in% c("DELAWARE", "PENNSYLVANIA", "WEST VIRGINIA")) |
           (aor == "New York City" & state == "NEW YORK") |
           (aor == "Buffalo" & state == "NEW YORK") |
           (aor == "Boston" & state %in% c("VERMONT", "NEW HAMPSHIRE", "MAINE", "MASSACHUSETTS", "CONNECTICUT", "RHODE ISLAND"))))
```

```{r}
invalid_state_aor_most_common <- invalid_state_aor %>%
  group_by(aor, state) %>%
  summarize(n = n()) %>%
  arrange(-n)
```